---
title: "Data Analyses Script (script 2)for the project, Association Between CAIDE Risk Score and Episodic Memory in Middle-Aged Females: The Impact of Natural Menopause"
author: "Julia Kearley, McGill University"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: flatly
    highlight: tango
    df_print: paged
    code_folding: hide
---

> **Note:** This script is for the data analyses (Script 2) to the manuscript:
>
> *Kearley, J. et al. (2025). Association Between CAIDE Risk Score and Episodic Memory in Middle-Aged Females: The Impact of Natural Menopause.* 
>

## Reproducibility Requirements

- **R version:** 4.3 or later  
- **Required R packages:** (see full list below)  
- **Data:** Cleaned dataset generated from `Kearley_bhamm_CAIDE_script1.rmd`

---

## Analysis Overview


### Part III: Descriptives: CAIDE Risk Score **(Figure 4)**  
- Descriptive statistics
- Visualizations

### Part IV: Main Analyses

#### Analysis 1: LMM — Task Predicting CS and CR, Stratified by Menopause Group **(Tables 3 & 4)**
Predictors: task, education, age.
- Predicting CS accuracy **(Model 1.2.a, Table 3)**   
- Predicting CR accuracy **(Model 1.2.b, Table 4)**   

#### Analysis 2: Group Difference in CAIDE Risk Score
- Mann-Whitney U test (pre- vs post-menopause)

#### Analysis 3: LMM — CAIDE Predicting CS and CR **(Tables 5 & 6)**
Separate models by menopause group:
- **Spatial Source Memory**
  - Premenopausal **(Model 2.1.a, Table 5)**  
  - Postmenopausal **(Model 2.1.b, Table 6)** 
- **Novelty Detection**
  - Premenopausal **(Model 2.2.a, Table 5)**  
  - Postmenopausal **(Model 2.2.b, Table 6)** y

### Part V: Post-Hoc Analyses **(Table 1)**
- Compare individual CAIDE risk categories between pre- and post-menopause groups


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)  # for portable file paths
``` 


## Load Packages
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(stats)
library(performance)
library(emmeans)
library(lme4)
library(rstatix)
library(car)
library(ggpubr)
library(stringr)
library(broom)
library(GGally)
library(tidyverse)
library(writexl)
library(interactions)
library(viridis)
library(sjmisc)
library(coin)
library(jtools)
library(reshape)
library(corrplot)
library(psych)

```

## Import Cleaned Data
```{r}
cleaned_data_df <- read.csv("script1_outputs/kearley_bhamm_CAIDE_cleaned_df_Jan_13_2025.csv")
cleaned_data_df$group_var <- factor(cleaned_data_df$group_var, levels = c("pre", "post"))
```
 Sample Size
```{r}
cleaned_data_df %>% count(group_var)
```

# Part III. Descriptives: CAIDE Risk score

### Visuals (Ordinal): CAIDE Risk Score Distribution (Figure 4)
```{r}
cleaned_data_df$CAIDE <- as.factor(cleaned_data_df$CAIDE)

p_CAIDE_meno <- ggplot(cleaned_data_df, aes(x = CAIDE, fill = group_var)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = c("#0072B2", "#D55E00"), name = "Menopause Group", labels = c("Premenopausal", "Postmenopausal")) +
  labs(x = "CAIDE Risk Score", y = "Count") +
  scale_y_continuous(breaks = seq(0, 12, 2), limits = c(0, 12)) +
  theme_pubclean() +
  theme(legend.position = "right")

p_CAIDE_meno
```


# Part IV. Main Analyses

## Analysis 1: Mixed-Effect Linear Regressions for Models Predicting Spatial Source Retrieval and Novelty Detection, Stratified by Menopause Group

### Spatial Source Retrieval (Model 1.2a, Table 3)
#### Premenopausal Group
```{r}
meno_cs_acc_pre <- cleaned_data_df %>%
  filter(group_var == "pre") %>%
  mutate(
    s2_age_std = scale(s2_age),
    edu_std = scale(edu)
  ) %>%
  select(id, s2_age, s2_age_std, edu, edu_std, cs_rate_easy, cs_rate_hard) %>%
  pivot_longer(cols = starts_with("cs_rate"), names_to = "task", values_to = "cs_rate") %>%
  mutate(
    id = factor(id),
    task = factor(str_replace(task, "cs_rate_", ""), levels = c("easy", "hard"))
  )

contrasts(meno_cs_acc_pre$task) <- contr.sum(2)
```

##### Fit Linear Mixed Models
```{r}
lmm1 <- lmerTest::lmer(cs_rate ~ task + s2_age_std + edu_std + (1|id), data = meno_cs_acc_pre, REML = FALSE)
lmm2 <- lmerTest::lmer(cs_rate ~ task * s2_age_std + edu_std + (1|id), data = meno_cs_acc_pre, REML = FALSE)
anova(lmm1, lmm2)

lmm_meno_cs_pre <- update(lmm1, REML = TRUE)
anova(lmm_meno_cs_pre, type = "III")
summary(lmm_meno_cs_pre)$coefficients
sjPlot::tab_model(lmm_meno_cs_pre, dv.labels = "Pre-Meno: Task, Age, Edu ~ CS", p.style = "scientific_stars", df.method = "satterthwaite")
check_collinearity(lmm_meno_cs_pre)
```

##### Estimated Marginal Means
```{r}
library(plyr)
out_EM_means_lmm_meno_cs_pre <- emmeans(lmm_meno_cs_pre, pairwise ~ task) %>% as.data.frame()
detach("package:plyr", unload = TRUE)
```

##### Visualizations
```{r}
library(jtools)
effect_plot(lmm_meno_cs_pre, pred = task, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  labs(x = "Task Difficulty", y = "Correct Source Accuracy") +
  theme_pubclean()
```

#### Postmenopausal Group
```{r}
meno_cs_acc_post <- cleaned_data_df %>%
  filter(group_var == "post") %>%
  mutate(
    s2_age_std = scale(s2_age),
    edu_std = scale(edu)
  ) %>%
  select(id, s2_age, s2_age_std, edu, edu_std, cs_rate_easy, cs_rate_hard) %>%
  pivot_longer(cols = starts_with("cs_rate"), names_to = "task", values_to = "cs_rate") %>%
  mutate(
    id = factor(id),
    task = factor(str_replace(task, "cs_rate_", ""), levels = c("easy", "hard"))
  )

contrasts(meno_cs_acc_post$task) <- contr.sum(2)
```

##### Fit Linear Mixed Models
```{r}
lmm1_post <- lmerTest::lmer(cs_rate ~ task + s2_age_std + edu_std + (1|id), data = meno_cs_acc_post, REML = FALSE)
lmm2_post <- lmerTest::lmer(cs_rate ~ task * s2_age_std + edu_std + (1|id), data = meno_cs_acc_post, REML = FALSE)
anova(lmm1_post, lmm2_post)

lmm_meno_cs_post <- update(lmm1_post, REML = TRUE)
anova(lmm_meno_cs_post, type = "III")
summary(lmm_meno_cs_post)$coefficients
sjPlot::tab_model(lmm_meno_cs_post, dv.labels = "Post-Meno: Task, Age, Edu ~ CS", p.style = "scientific_stars", df.method = "satterthwaite")
check_collinearity(lmm_meno_cs_post)
```


##### Visualizations
```{r}
# Task Effect
plot_task <- effect_plot(lmm_meno_cs_post, pred = task, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  labs(x = "Task Difficulty", y = "Correct Source Accuracy") +
  theme_pubclean()

# Age Effect
plot_age <- effect_plot(lmm_meno_cs_post, pred = s2_age_std, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  labs(x = "Age (Standardized)", y = "Correct Source Accuracy") +
  theme_pubclean()

# Education Effect
plot_edu <- effect_plot(lmm_meno_cs_post, pred = edu_std, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  labs(x = "Education (Standardized)", y = "Correct Source Accuracy") +
  theme_pubclean()

list(plot_task, plot_age, plot_edu)
```

### Novelty Detection (Model 1.2b, Table 4 in manuscript)
#### Premenopausal Group
```{r}
meno_cr_acc_pre <- cleaned_data_df %>%
  filter(group_var == "pre") %>%
  mutate(
    s2_age_std = scale(s2_age),
    edu_std = scale(edu)
  ) %>%
  select(id, s2_age, s2_age_std, edu, edu_std, cr_rate_easy, cr_rate_hard) %>%
  pivot_longer(cols = starts_with("cr_rate"), names_to = "task", values_to = "cr_rate") %>%
  mutate(
    id = factor(id),
    task = factor(str_replace(task, "cr_rate_", ""), levels = c("easy", "hard"))
  )

contrasts(meno_cr_acc_pre$task) <- contr.sum(2)
```

##### Model Fitting
```{r}
lmm1_cr_pre <- lmerTest::lmer(cr_rate ~ task + s2_age_std + edu_std + (1|id), data = meno_cr_acc_pre, REML = FALSE)
lmm2_cr_pre <- lmerTest::lmer(cr_rate ~ task * s2_age_std + edu_std + (1|id), data = meno_cr_acc_pre, REML = FALSE)
anova(lmm1_cr_pre, lmm2_cr_pre)

lmm_meno_cr_pre <- update(lmm1_cr_pre, REML = TRUE)
anova(lmm_meno_cr_pre, type = "III")
summary(lmm_meno_cr_pre)$coefficients
sjPlot::tab_model(lmm_meno_cr_pre, dv.labels = "Pre-Meno: Task, Age, Edu ~ CR", p.style = "scientific_stars", df.method = "satterthwaite")
check_collinearity(lmm_meno_cr_pre)
```

##### Visualization
```{r}
effect_plot(model = lmm_meno_cr_pre, pred = task, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7, x.label = "Task Difficulty", y.label = "Correct Rejection Accuracy") +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  theme_pubclean()
```

#### Postmenopausal Group
```{r}
meno_cr_acc_post <- cleaned_data_df %>%
  filter(group_var == "post") %>%
  mutate(
    s2_age_std = scale(s2_age),
    edu_std = scale(edu)
  ) %>%
  select(id, s2_age, s2_age_std, edu, edu_std, cr_rate_easy, cr_rate_hard) %>%
  pivot_longer(cols = starts_with("cr_rate"), names_to = "task", values_to = "cr_rate") %>%
  mutate(
    id = factor(id),
    task = factor(str_replace(task, "cr_rate_", ""), levels = c("easy", "hard"))
  )

contrasts(meno_cr_acc_post$task) <- contr.sum(2)
```

##### Model Fitting
```{r}
lmm1_cr_post <- lmerTest::lmer(cr_rate ~ task + s2_age_std + edu_std + (1|id), data = meno_cr_acc_post, REML = FALSE)
lmm2_cr_post <- lmerTest::lmer(cr_rate ~ task * s2_age_std + edu_std + (1|id), data = meno_cr_acc_post, REML = FALSE)
anova(lmm1_cr_post, lmm2_cr_post)

lmm_meno_cr_post <- update(lmm1_cr_post, REML = TRUE)
anova(lmm_meno_cr_post, type = "III")
summary(lmm_meno_cr_post)$coefficients
sjPlot::tab_model(lmm_meno_cr_post, dv.labels = "Post-Meno: Task, Age, Edu ~ CR", p.style = "scientific_stars", df.method = "satterthwaite")
check_collinearity(lmm_meno_cr_post)
```

##### Visualization
```{r}
effect_plot(model = lmm_meno_cr_post, pred = task, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7, x.label = "Task Difficulty", y.label = "Correct Rejection Accuracy") +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  theme_pubclean()
```

## Analyis 2: CAIDE Score between menopause groups
### Conduct Mann-Whitney U test to examine differences in CAIDE score by menopause status.
```{r}
wilcox.test(as.numeric(CAIDE) ~ group_var, data = cleaned_data_df, exact=FALSE)
```


## Analysis 3: CAIDE Score Predicting Spatial Source Retrieval and Novelty Detection, Stratified by Menopause 
### Spatial Source Memory (CS) ~ CAIDE in Premenopausal Females (Model 2.1a, Table 5)
Prepare data:
```{r}
cs_acc_CAIDE_pre <- cleaned_data_df %>%
  filter(group_var == "pre") %>%
  select(id, group_var, CAIDE, cs_rate_easy, cs_rate_hard) %>%
  mutate(CAIDE = factor(CAIDE)) %>%
  pivot_longer(cols = starts_with("cs_rate"), names_to = "task", values_to = "cs_rate") %>%
  mutate(task = str_replace(task, "cs_rate_", "")) %>%
  mutate(task = factor(task, levels = c("easy", "hard")))
```

Reverse Helmert contrast coding:
```{r}
cs_acc_CAIDE_pre <- cs_acc_CAIDE_pre %>%
  mutate(CAIDE = factor(CAIDE, levels = c("0","1","2","3","4","5","6","7")))

  # create contrasts
    r_helmert_1 = matrix(c(-1/2, 1/2, 0, 0, 0, 0, 0,0,
                            -1/3, -1/3, 2/3, 0, 0, 0, 0,0,
                            -1/4, -1/4, -1/4, 3/4, 0, 0, 0,0,
                            -1/5, -1/5, -1/5, -1/5, 4/5, 0, 0,0,
                            -1/6, -1/6, -1/6, -1/6, -1/6, 5/6, 0,0,
                            -1/7, -1/7, -1/7, -1/7, -1/7, -1/7, 6/7,0,
                           -1/8, -1/8, -1/8, -1/8, -1/8, -1/8, -1/8, 7/8
                            ), ncol = 7)

    r_helmert_1

contrasts(cs_acc_CAIDE_pre$task) <- contr.sum(2)
contrasts(cs_acc_CAIDE_pre$task)

contrasts(cs_acc_CAIDE_pre$CAIDE) <- r_helmert_1
contrasts(cs_acc_CAIDE_pre$CAIDE)
```

Fit model:
```{r}
lmm_CAIDE_cs_pre <- lmerTest::lmer(cs_rate ~ task + CAIDE + (1|id), data = cs_acc_CAIDE_pre, REML = TRUE)
```

Outputs:
```{r}
anova(lmm_CAIDE_cs_pre, type = "III")
summary(lmm_CAIDE_cs_pre)$coefficients
sjPlot::tab_model(
  lmm_CAIDE_cs_pre,
  dv.labels = "Pre-Meno: Task and CAIDE Risk score predicting CS",
  p.style = "scientific_stars",
  df.method = "satterthwaite"
)

check_collinearity(lmm_CAIDE_cs_pre)
check_outliers(lmm_CAIDE_cs_pre)
```

Visualizations:
```{r}
effect_plot(model = lmm_CAIDE_cs_pre, pred = task, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  labs(x = "Task Difficulty", y = "Correct Source Accuracy") +
  theme_pubclean()

effect_plot(model = lmm_CAIDE_cs_pre, pred = CAIDE, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  labs(x = "CAIDE", y = "Correct Source Accuracy") +
  theme_pubclean()
```


### Spatial Source Memory (CS) ~ CAIDE in Postmenopausal Females (Model 2.1b, Table 6)

Prepare data:
```{r}
cs_acc_CAIDE_post <- cleaned_data_df %>%
  filter(group_var == "post") %>%
  select(id, group_var, CAIDE, cs_rate_easy, cs_rate_hard) %>%
  mutate(CAIDE = factor(CAIDE)) %>%
  pivot_longer(cols = starts_with("cs_rate"), names_to = "task", values_to = "cs_rate") %>%
  mutate(task = str_replace(task, "cs_rate_", "")) %>%
  mutate(task = factor(task, levels = c("easy", "hard")))
```

Reverse Helmert contrast coding:
```{r}
cs_acc_CAIDE_post <- mutate(cs_acc_CAIDE_post, CAIDE = factor(CAIDE, levels = c("4", "5", "6", "7", "8", "9", "10")))

r_helmert_1 = matrix(c(-1/2, 1/2, 0, 0, 0, 0, 0,
                            -1/3, -1/3, 2/3, 0, 0, 0, 0,
                            -1/4, -1/4, -1/4, 3/4, 0, 0, 0,
                            -1/5, -1/5, -1/5, -1/5, 4/5, 0, 0,
                            -1/6, -1/6, -1/6, -1/6, -1/6, 5/6, 0,
                            -1/7, -1/7, -1/7, -1/7, -1/7, -1/7, 6/7
                            ), ncol = 6)

    r_helmert_1

contrasts(cs_acc_CAIDE_post$task) <- contr.sum(2)
contrasts(cs_acc_CAIDE_post$CAIDE) <- r_helmert_1
```

Fit model:
```{r}
lmm_CAIDE_cs_post <- lmerTest::lmer(cs_rate ~ task + CAIDE + (1|id), data = cs_acc_CAIDE_post, REML = TRUE)
```

Outputs:
```{r}
anova(lmm_CAIDE_cs_post, type = "III")
summary(lmm_CAIDE_cs_post)$coefficients
sjPlot::tab_model(
  lmm_CAIDE_cs_post,
  dv.labels = "Post-Meno: Task and CAIDE Risk score predicting CS",
  p.style = "scientific_stars",
  df.method = "satterthwaite"
)

check_collinearity(lmm_CAIDE_cs_post)
check_outliers(lmm_CAIDE_cs_post)
```

Visualizations:
```{r}
effect_plot(model = lmm_CAIDE_cs_post, pred = task, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  labs(x = "Task Difficulty", y = "Correct Source Accuracy") +
  theme_pubclean()

effect_plot(model = lmm_CAIDE_cs_post, pred = CAIDE, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  labs(x = "CAIDE", y = "Correct Source Accuracy") +
  theme_pubclean()
```


### Novelty Detection (CR) ~ CAIDE in Premenopausal Females ( Model 2.2a, Table 5)
Prepare data:
```{r}
cr_acc_CAIDE_pre <- cleaned_data_df %>%
  filter(group_var == "pre") %>%
  select(id, group_var, CAIDE, cr_rate_easy, cr_rate_hard) %>%
  mutate(CAIDE = factor(CAIDE)) %>%
  pivot_longer(cols = starts_with("cr_rate"), names_to = "task", values_to = "cr_rate") %>%
  mutate(task = str_replace(task, "cr_rate_", "")) %>%
  mutate(task = factor(task, levels = c("easy", "hard")))
```

Reverse Helmert contrast coding:
```{r}
cr_acc_CAIDE_pre <- mutate(cr_acc_CAIDE_pre, CAIDE = factor(CAIDE, levels = c("0","1","2","3","4","5","6","7")))

  # create contrasts
    r_helmert_2 = matrix(c(-1/2, 1/2, 0, 0, 0, 0, 0,0,
                            -1/3, -1/3, 2/3, 0, 0, 0, 0,0,
                            -1/4, -1/4, -1/4, 3/4, 0, 0, 0,0,
                            -1/5, -1/5, -1/5, -1/5, 4/5, 0, 0,0,
                            -1/6, -1/6, -1/6, -1/6, -1/6, 5/6, 0,0,
                            -1/7, -1/7, -1/7, -1/7, -1/7, -1/7, 6/7,0,
                           -1/8, -1/8, -1/8, -1/8, -1/8, -1/8, -1/8, 7/8
                            ), ncol = 7)

    r_helmert_2
    
contrasts(cr_acc_CAIDE_pre$task) <- contr.sum(2)
contrasts(cr_acc_CAIDE_pre$CAIDE) <- r_helmert_2
```

Fit model:
```{r}
lmm_CAIDE_cr_pre <- lmerTest::lmer(cr_rate ~ task + CAIDE + (1|id), data = cr_acc_CAIDE_pre, REML = TRUE)
```

Outputs:
```{r}
anova(lmm_CAIDE_cr_pre, type = "III")
summary(lmm_CAIDE_cr_pre)$coefficients
sjPlot::tab_model(
  lmm_CAIDE_cr_pre,
  dv.labels = "Pre-Meno: Task and CAIDE Risk score predicting CR",
  p.style = "scientific_stars",
  df.method = "satterthwaite"
)
```

Visualizations:
```{r}
effect_plot(model = lmm_CAIDE_cr_pre, pred = task, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  labs(x = "Task Difficulty", y = "Correct Rejection Accuracy") +
  theme_pubclean()

effect_plot(model = lmm_CAIDE_cr_pre, pred = CAIDE, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  labs(x = "CAIDE", y = "Correct Rejection Accuracy") +
  theme_pubclean()
```

### Novelty Detection (CR) ~ CAIDE in Posmenopausal Females (Model 2.2b, Table 6)

Prepare data:
```{r}
cr_acc_CAIDE_post <- cleaned_data_df %>%
  filter(group_var == "post") %>%
  select(id, group_var, CAIDE, cr_rate_easy, cr_rate_hard) %>%
  mutate(CAIDE = factor(CAIDE)) %>%
  pivot_longer(cols = starts_with("cr_rate"), names_to = "task", values_to = "cr_rate") %>%
  mutate(task = str_replace(task, "cr_rate_", "")) %>%
  mutate(task = factor(task, levels = c("easy", "hard")))
```

Reverse Helmert contrast coding:
```{r}
cr_acc_CAIDE_post <- mutate(cr_acc_CAIDE_post, CAIDE = factor(CAIDE, levels = c("4", "5", "6", "7", "8", "9", "10")))

# create contrasts
    r_helmert_2 = matrix(c(-1/2, 1/2, 0, 0, 0, 0, 0,
                            -1/3, -1/3, 2/3, 0, 0, 0, 0,
                            -1/4, -1/4, -1/4, 3/4, 0, 0, 0,
                            -1/5, -1/5, -1/5, -1/5, 4/5, 0, 0,
                            -1/6, -1/6, -1/6, -1/6, -1/6, 5/6, 0,
                            -1/7, -1/7, -1/7, -1/7, -1/7, -1/7, 6/7
                            ), ncol = 6)

    r_helmert_2

contrasts(cr_acc_CAIDE_post$task) <- contr.sum(2)
contrasts(cr_acc_CAIDE_post$CAIDE) <- r_helmert_2
```

Fit model:
```{r}
lmm_CAIDE_cr_post <- lmerTest::lmer(cr_rate ~ task + CAIDE + (1|id), data = cr_acc_CAIDE_post, REML = TRUE)
```

Outputs:
```{r}
anova(lmm_CAIDE_cr_post, type = "III")
summary(lmm_CAIDE_cr_post)$coefficients
sjPlot::tab_model(
  lmm_CAIDE_cr_post,
  dv.labels = "Post-Meno: Task and CAIDE Risk score predicting CR",
  p.style = "scientific_stars",
  df.method = "satterthwaite"
)
```

Visualizations:
```{r}
effect_plot(model = lmm_CAIDE_cr_post, pred = task, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  labs(x = "Task Difficulty", y = "Correct Rejection Accuracy") +
  theme_pubclean()

effect_plot(model = lmm_CAIDE_cr_post, pred = CAIDE, plot.points = TRUE, jitter = 0.02, line.thickness = 0.7) +
  labs(x = "CAIDE", y = "Correct Rejection Accuracy") +
  theme_pubclean()

```



# Part V: Post-Hocs on individual CAIDE risk score variables (Table 1)

## Risk Bin Distirbution for each CAIDE Variable 

Prepare long format dataframe:
```{r}
x_vars <- c("Age_coded", "BMI_coded", "Systolic_coded", "Exercise_coded", "Chol_coded", "Edu_coded")

risk_factors_df_long_2 <- cleaned_data_df %>%
  select(id, group_var, all_of(x_vars)) %>%
  mutate(across(all_of(x_vars), as.factor)) %>%
  pivot_longer(cols = all_of(x_vars), names_to = "risk_factor", values_to = "value") %>% 
  convert_as_factor(id, group_var)
```

Descriptive statistics:
```{r}
# Counts and frequencies, overall
risk_factors_df_long_2 %>%
  count(risk_factor, value) %>%
  group_by(risk_factor) %>%
  mutate(freq = n / sum(n))

# Count and frequencies, by menopause group
out_summary_CAIDE_meno <- risk_factors_df_long_2 %>%
  count(group_var, risk_factor, value) %>%
  group_by(group_var, risk_factor) %>%
  mutate(freq = round(n / sum(n), 2)) %>%
  ungroup() %>%
  pivot_wider(
    names_from = group_var,
    values_from = c(n, freq),
    names_sep = "_"
  ) %>%
  select(risk_factor, value, everything())
out_summary_CAIDE_meno

```

Statistical tests:
```{r}
dat <- cleaned_data_df %>% select(group_var, all_of(x_vars))

kruskal_test <- lapply(dat[-1], function(x) kruskal.test(dat$group_var, x))
do.call(rbind, kruskal_test)[, c(1, 3)]

mann_test <- lapply(dat[-1], function(x) wilcox.test(x ~ dat$group_var, exact = FALSE))
do.call(rbind, mann_test)[, c(1, 3)]
```
Statistical tests summary table:
```{r}
dat <- cleaned_data_df %>% select(group_var, all_of(x_vars))

kruskal_test <- lapply(dat[-1], function(x) kruskal.test(dat$group_var, x))
kruskal_df <- data.frame(
  Variable = names(dat)[-1],
  Test = "Kruskal-Wallis",
  Statistic = sapply(kruskal_test, function(x) round(x$statistic, 3)),
  p_value = sapply(kruskal_test, function(x) format.pval(x$p.value, digits = 4, eps = .0001))
)

mann_test <- lapply(dat[-1], function(x) wilcox.test(x ~ dat$group_var, exact = FALSE))
mann_df <- data.frame(
  Variable = names(dat)[-1],
  Test = "Mann-Whitney U",
  Statistic = sapply(mann_test, function(x) round(x$statistic, 3)),
  p_value = sapply(mann_test, function(x) format.pval(x$p.value, digits = 4, eps = .0001))
)

test_results_df <- bind_rows(kruskal_df, mann_df)
test_results_df
```

Visualizations:
```{r}
dat <- cleaned_data_df %>%
  select(id, group_var, all_of(x_vars)) %>%
  convert_as_factor(!!!syms(c("id", "group_var", x_vars)))

plot_bar <- function(var, title, labels) {
  ggplot(dat, aes_string(x = var)) +
    geom_bar(aes(fill = group_var), position = "dodge") +
    scale_fill_manual(name = "Menopause Group", values = c("#0072B2", "#D55E00"), labels = c("Premenopausal", "Postmenopausal")) +
    scale_x_discrete(labels = labels) +
    scale_y_continuous(breaks = seq(0, 50, by = 10), limits = c(0, 50)) +
    labs(x = "", y = "Count", title = title) +
    theme_pubclean() +
    theme(plot.title = element_text(hjust = 0.5))
}

age_p <- plot_bar("Age_coded", "Age", c("< 47 yrs", "47-53 yrs", "> 53 yrs"))
edu_p <- plot_bar("Edu_coded", "Education", c("< 12 yrs", "12-16 yrs", "> 16 yrs"))
bmi_p <- plot_bar("BMI_coded", "BMI", c("<= 30", "> 30"))
chol_p <- plot_bar("Chol_coded", "Cholesterol", c("<= 6.2", "> 6.2"))
bp_p <- plot_bar("Systolic_coded", "Systolic BP", c("<= 140", "> 140"))
exercise_p <- plot_bar("Exercise_coded", "Exercise", c("Active", "Inactive"))

p_vrf2 <- ggarrange(age_p, edu_p, bmi_p, chol_p, bp_p, exercise_p, 
                    ncol = 3, nrow = 2, common.legend = TRUE, legend = "bottom")
p_vrf2
```


### Save Output
```{r}
# # Save plots
# plots <- list(
#   p_meno_cs_age_raw = p_meno_cs_age_raw,
#   p_meno_cs_acc_raw = p_meno_cs_acc_raw,
#   p_meno_cr_age_raw = p_meno_cr_age_raw,
#   p_meno_cr_acc_raw = p_meno_cr_acc_raw,
#   p_CAIDE_meno = p_CAIDE_meno,
#   p_vrf2 = p_vrf2
# )

# lapply(names(plots), function(n) {
#   ggsave(here("script2_outputs", sprintf("%s.jpeg", n)), plots[[n]], width = 8, height = 8, dpi = 700)
# })

# #Save model objects
# save(list = c(
#   "cleaned_data_df", "meno_cs_acc", "meno_cr_acc",
#   "cs_acc_CAIDE_pre", "cs_acc_CAIDE_post",
#   "cr_acc_CAIDE_pre", "cr_acc_CAIDE_post",
#   "lmm_CAIDE_cs_pre", "lmm_CAIDE_cs_post",
#   "lmm_CAIDE_cr_pre", "lmm_CAIDE_cr_post"
# ), file = here("script2_outputs", "supplemental_objects_Jan13_2025"))

```

