---
title: "Data Analyses Script (Supplemental Materials) for the project, Association Between CAIDE Risk Score and Episodic Memory in Middle-Aged Females: The Impact of Natural Menopause"
author: "Julia Kearley, McGill University"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
    theme: flatly
    highlight: tango
    df_print: paged
    code_folding: hide
---

> **Note:** This script is for the **Supplemental Materials Analyses** to the manuscript:
>
> *Kearley, J. et al. (2025). Association Between CAIDE Risk Score and Episodic Memory in Middle-Aged Females: The Impact of Natural Menopause.*

## Reproducibility Requirements

- **R version:** 4.3 or later
- **Required R packages:** (see loaded libraries below)
- **Data:** Saved R objects from `Kearley_bhamm_memory_rf_script2.rmd`

---

## Analysis Overview

### Sup1 (Figure S2): Modified CAIDE Score (Excluding Age Points)
- Summary statistics (overall and by group)
- Mann-Whitney U test
- Histogram by menopause status

### Sup2 (Table S1): LRT Tests (CS retrieval ~ Task × CAIDE)
- Premenopausal and postmenopausal groups

### Sup3 (Table S2): LRT Tests (Novelty detection ~ Task × CAIDE)
- Premenopausal and postmenopausal groups

### Sup4 (Table S3): Between-Group Mixed-Effect Linear Regressions — CS and CR (Models 1.1a & 1.1b)
- Task × group_var interaction and marginal effects

### Sup5 (Table S4): Stepwise Regression — CS
- Predictors: task, menopause group, age, and education
- Final model based on AIC selection

### Sup6 (Table S5): Stepwise Regression — CR
- Same procedure as Sup5 for novelty detection

### Sup7 (Table S6): Post Hoc — Interaction Between Age at FMP and CAIDE
- Investigates age at final menstrual period (FMP) as moderator of CAIDE × CS

### Sup8 (Table S7): Mediation Analysis — CAIDE as Mediator
- Tests if CAIDE mediates menopause status → CS accuracy
- Manual and lavaan SEM approach
- Includes binned CAIDE and bootstrapping

### Sup9 (Table S8): Mixed-Effects Models — CAIDE Continuous Predicting CS
- Separate models for pre- and post-menopause groups
- Task and CAIDE predicting CS accuracy


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
```

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(stats)
library(performance)
library(emmeans)
library(lme4)
library(rstatix)
library(car)
library(ggpubr)
library(stringr)
library(broom)
library(GGally)
library(tidyverse)
library(writexl)
library(interactions)
library(jtools)
```

```{r}
apatheme1 <- theme_bw() +
  theme(panel.grid = element_blank(),
        axis.line = element_line(),
        text = element_text(family='Times', size = 20),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5, size = 30))

apatheme2 <- theme_bw() +
  theme(panel.grid.major = element_blank(),
        text = element_text(family='Times', size = 20),
        plot.title = element_text(hjust = 0.5, size = 30))
```

```{r}
load("supplemental_objects_Jan13_2025")
```

## Sup1 (Figure S2): Modified CAIDE Score (Excluding Age Points)

```{r}
cleaned_data_df %>%
  summarise(mean = round(mean(CAIDE_no_age), 2),
            sd = round(sd(CAIDE_no_age), 2),
            min = min(CAIDE_no_age),
            max = max(CAIDE_no_age))

cleaned_data_df %>%
  group_by(group_var) %>%
  summarise(mean = round(mean(CAIDE_no_age), 2),
            sd = round(sd(CAIDE_no_age), 2),
            min = min(CAIDE_no_age),
            max = max(CAIDE_no_age))

cleaned_data_df %>%
  count(CAIDE_no_age, group_var) %>%
  mutate(Freq = n / sum(n))

wilcox.test(as.numeric(CAIDE_no_age) ~ group_var, data = cleaned_data_df, exact=FALSE)

cleaned_data_df$CAIDE_no_age <- as.factor(cleaned_data_df$CAIDE_no_age)

sup_caide_p <- ggplot(cleaned_data_df, aes(x = CAIDE_no_age, fill = group_var)) +
  geom_histogram(position = "dodge", stat = "count") +
  scale_fill_manual(name = "Menopause Group",
                    labels = c("Premenopausal", "Postmenopausal"),
                    values = c("#0072B2", "#D55E00")) +
  scale_y_continuous(limits = c(0, 12), breaks = seq(0, 12, 2)) +
  labs(x = "Modified CAIDE Risk Score\n(No age points)", y = "Count") +
  theme_pubclean() +
  theme(legend.position = "right")

sup_caide_p
```

## Sup2 (Table S1): LRT for Models Predicting CS from CAIDE score (Analysis #3 in paper) — Premenopausal

```{r}
contrasts(cs_acc_CAIDE_pre$task) <- contr.sum(2)

lmm_CAIDE_cs_pre_0 <- lmer(cs_rate ~ task + (1|id), data = cs_acc_CAIDE_pre, REML = FALSE)
lmm_CAIDE_cs_pre_1 <- lmer(cs_rate ~ task + CAIDE + (1|id), data = cs_acc_CAIDE_pre, REML = FALSE)
lmm_CAIDE_cs_pre_2 <- lmer(cs_rate ~ task * CAIDE + (1|id), data = cs_acc_CAIDE_pre, REML = FALSE)

lrt_pre <- anova(lmm_CAIDE_cs_pre_0, lmm_CAIDE_cs_pre_1, lmm_CAIDE_cs_pre_2)
as.data.frame(lrt_pre)

library(MuMIn)
r.squaredGLMM(lmm_CAIDE_cs_pre_0)
r.squaredGLMM(lmm_CAIDE_cs_pre_1)
r.squaredGLMM(lmm_CAIDE_cs_pre_2)
```

## Sup2 (continued): Postmenopausal Group

```{r}
contrasts(cs_acc_CAIDE_post$task) <- contr.sum(2)

lmm_CAIDE_cs_post_0 <- lmer(cs_rate ~ task + (1|id), data = cs_acc_CAIDE_post, REML = FALSE)
lmm_CAIDE_cs_post_1 <- lmer(cs_rate ~ task + CAIDE + (1|id), data = cs_acc_CAIDE_post, REML = FALSE)
lmm_CAIDE_cs_post_2 <- lmer(cs_rate ~ task * CAIDE + (1|id), data = cs_acc_CAIDE_post, REML = FALSE)

lrt_post <- anova(lmm_CAIDE_cs_post_0, lmm_CAIDE_cs_post_1, lmm_CAIDE_cs_post_2)
as.data.frame(lrt_post)
```

## Sup3 (Table S2): Novelty Detection Models — Premenopausal

```{r}
contrasts(cr_acc_CAIDE_pre$task) <- contr.sum(2)

lmm_CAIDE_cr_pre_0 <- lmer(cr_rate ~ task + (1|id), data = cr_acc_CAIDE_pre, REML = FALSE)
lmm_CAIDE_cr_pre_1 <- lmer(cr_rate ~ task + CAIDE + (1|id), data = cr_acc_CAIDE_pre, REML = FALSE)
lmm_CAIDE_cr_pre_2 <- lmer(cr_rate ~ task * CAIDE + (1|id), data = cr_acc_CAIDE_pre, REML = FALSE)

anova(lmm_CAIDE_cr_pre_0, lmm_CAIDE_cr_pre_1, lmm_CAIDE_cr_pre_2)
```

## Sup3 (continued): Postmenopausal Group

```{r}
contrasts(cr_acc_CAIDE_post$task) <- contr.sum(2)

lmm_CAIDE_cr_post_0 <- lmer(cr_rate ~ task + (1|id), data = cr_acc_CAIDE_post, REML = FALSE)
lmm_CAIDE_cr_post_1 <- lmer(cr_rate ~ task + CAIDE + (1|id), data = cr_acc_CAIDE_post, REML = FALSE)
lmm_CAIDE_cr_post_2 <- lmer(cr_rate ~ task * CAIDE + (1|id), data = cr_acc_CAIDE_post, REML = FALSE)

anova(lmm_CAIDE_cr_post_0, lmm_CAIDE_cr_post_1, lmm_CAIDE_cr_post_2)
```
## Sup4 (Table S3): Between-Group Mixed-Effect Linear Regressions — CS and CR

### Model 1.1a: CS Accuracy
```{r}
sup_4_lmm_meno_cs <- lmer(cs_rate ~ task * group_var + edu_std + (1|id), data = meno_cs_acc, REML = TRUE)
anova(sup_4_lmm_meno_cs, type = "III")
summary(sup_4_lmm_meno_cs)$coefficients

sjPlot::tab_model(sup_4_lmm_meno_cs,
                  dv.labels = "Task and Menopause Status predicting CS", p.style = "scientific_stars", df.method = "satterthwaite")
check_collinearity(sup_4_lmm_meno_cs)

sup_4_out_EM_means_lmm_meno_cs <- bind_rows(
  as.data.frame(emmeans(sup_4_lmm_meno_cs, pairwise ~ task)),
  as.data.frame(emmeans(sup_4_lmm_meno_cs, pairwise ~ group_var))
)
sup_4_out_EM_means_lmm_meno_cs

sup_4_out_EM_contrasts_lmm_meno_cs <- as.data.frame(emmeans(sup_4_lmm_meno_cs, pairwise ~ group_var | task))
sup_4_out_EM_contrasts_lmm_meno_cs
```

#### Plot: Main Effects — CS
```{r}
effect_plot(sup_4_lmm_meno_cs, pred = group_var, plot.points = TRUE, jitter = 0.02,
            line.thickness = 0.7, x.label = "Menopause Group", y.label = "Correct Source Accuracy") +
  scale_x_discrete(labels = c("Premenopausal", "Postmenopausal")) +
  theme_pubclean()

effect_plot(sup_4_lmm_meno_cs, pred = task, plot.points = TRUE, jitter = 0.02,
            line.thickness = 0.7, x.label = "Task Difficulty", y.label = "Correct Source Accuracy") +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  theme_pubclean()
```

### Model 1.1b: CR Accuracy
```{r}
sup_4_lmm_meno_cr <- lmer(cr_rate ~ task * group_var + edu_std + (1|id), data = meno_cr_acc, REML = TRUE)
anova(sup_4_lmm_meno_cr, type = "III")
summary(sup_4_lmm_meno_cr)$coefficients

sjPlot::tab_model(sup_4_lmm_meno_cr,
                  dv.labels = "Task and Menopause Status Predicting CR", p.style = "scientific_stars")
check_collinearity(sup_4_lmm_meno_cr)

sup_4_out_EM_means_lmm_meno_cr <- bind_rows(
  as.data.frame(emmeans(sup_4_lmm_meno_cr, pairwise ~ task)),
  as.data.frame(emmeans(sup_4_lmm_meno_cr, pairwise ~ group_var))
)
sup_4_out_EM_means_lmm_meno_cr

sup_4_out_EM_contrasts_lmm_meno_cr <- as.data.frame(emmeans(sup_4_lmm_meno_cr, pairwise ~ group_var | task))
sup_4_out_EM_contrasts_lmm_meno_cr
```

#### Plot: Main Effects — CR
```{r}
effect_plot(sup_4_lmm_meno_cr, pred = group_var, plot.points = TRUE, jitter = 0.02,
            line.thickness = 0.7, x.label = "Menopause Group", y.label = "Correct Rejection Accuracy") +
  scale_x_discrete(labels = c("Premenopausal", "Postmenopausal")) +
  theme_pubclean()

effect_plot(sup_4_lmm_meno_cr, pred = task, plot.points = TRUE, jitter = 0.02,
            line.thickness = 0.7, x.label = "Task Difficulty", y.label = "Correct Rejection Accuracy") +
  scale_x_discrete(labels = c("Easy", "Hard")) +
  theme_pubclean()
```




## Sup5 (Table S4): Stepwise Regression — CS
```{r}
full.model_cs <- lmer(cs_rate ~ task * group_var * s2_age_std + edu_std + (1|id), data = meno_cs_acc, REML=TRUE)
fixmodel_cs <- lm(formula(full.model_cs, fixed.only=TRUE), data=eval(getCall(full.model_cs)$data))
step(fixmodel_cs)

stepwise_cs <- lmer(cs_rate ~ task + group_var + s2_age_std + edu_std + task:group_var + (1|id), data = meno_cs_acc, REML=TRUE)
sjPlot::tab_model(stepwise_cs, dv.labels = "Stepwise model predicting CS", p.style = "scientific_stars", df.method = "satterthwaite")
```

## Sup6 (Table S5): Stepwise Regression — CR
```{r}
full.model_cr <- lmer(cr_rate ~ task * group_var * s2_age_std + edu_std + (1|id), data = meno_cr_acc, REML=TRUE)
fixmodel_cr <- lm(formula(full.model_cr, fixed.only=TRUE), data=eval(getCall(full.model_cr)$data))
step(fixmodel_cr)

stepwise_cr <- lmer(cr_rate ~ task + s2_age_std + (1|id), data = meno_cr_acc, REML=TRUE)
sjPlot::tab_model(stepwise_cr, dv.labels = "Stepwise model predicting CR", p.style = "scientific_stars", df.method = "satterthwaite")
```

## Sup7 (Table S6): Post Hoc — Interaction Between Age at FMP and CAIDE (CS)

```{r}
age_FMP <- cleaned_data_df %>%
  select(id, group_var, CAIDE, edu, s2_age, s2_fmp_mths) %>%
  filter(group_var == "post") %>%
  mutate(age_fmp = s2_age - s2_fmp_mths / 12)

cor.test(age_FMP$s2_age, age_FMP$age_fmp)
cor.test(as.numeric(age_FMP$CAIDE), age_FMP$age_fmp)
```

Create df: Merge FMP data with cs dataframe
```{r}
x <- age_FMP %>% select(id, s2_age, edu, age_fmp)
fmp_age_cs_df <- merge(x, cs_acc_CAIDE_post, by = "id",all=FALSE)
fmp_age_cs_df$s2_age_bin_2 <- as.numeric(cut_number(fmp_age_cs_df$s2_age,3))
```

```{r}
age_fmp_CS_plot <- ggplot(fmp_age_cs_df, aes(x = age_fmp, y = cs_rate, color = as.factor(s2_age_bin_2))) +
  facet_grid(~task) +
  geom_point() +
  geom_smooth(method = lm) +
  scale_y_continuous(name = "CS Rate", limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  scale_x_continuous(name = "FMP (Age)") +
  scale_colour_manual(values = c("green4", "orange", "red3"), labels = c("Below -1 SD", "Within ±1 SD", "Above +1 SD")) +
  labs(colour = "Age") +
  theme_pubclean() +
  theme(legend.position = "bottom")

age_fmp_CS_plot
```

```{r}
cs_acc_CAIDE_fmp_post <- merge(cs_acc_CAIDE_post, age_FMP %>% select(id, age_fmp), by = "id") %>%
  na.omit() %>%
  mutate(CAIDE = factor(CAIDE, levels = c("4","5","6","7","8","9","10")))

r_helmert_1 <- matrix(c(
  -1/2, 1/2, 0, 0, 0, 0, 0,
  -1/3, -1/3, 2/3, 0, 0, 0, 0,
  -1/4, -1/4, -1/4, 3/4, 0, 0, 0,
  -1/5, -1/5, -1/5, -1/5, 4/5, 0, 0,
  -1/6, -1/6, -1/6, -1/6, -1/6, 5/6, 0,
  -1/7, -1/7, -1/7, -1/7, -1/7, -1/7, 6/7
), ncol = 6)

contrasts(cs_acc_CAIDE_fmp_post$task) <- contr.sum(2)
contrasts(cs_acc_CAIDE_fmp_post$CAIDE) <- r_helmert_1
```

```{r}
lmm_CAIDE_cs_post_fmp0 <- lmer(cs_rate ~ task + CAIDE + (1|id), data = cs_acc_CAIDE_fmp_post, REML=FALSE)
lmm_CAIDE_cs_post_fmp1 <- lmer(cs_rate ~ task + CAIDE + age_fmp + (1|id), data = cs_acc_CAIDE_fmp_post, REML=FALSE)
lmm_CAIDE_cs_post_fmp2 <- lmer(cs_rate ~ task + CAIDE * age_fmp + (1|id), data = cs_acc_CAIDE_fmp_post, REML=FALSE)

anova(lmm_CAIDE_cs_post_fmp0, lmm_CAIDE_cs_post_fmp1, lmm_CAIDE_cs_post_fmp2)
```

```{r}
lmm_CAIDE_cs_post_fmp2 <- lmer(cs_rate ~ task + CAIDE * age_fmp + (1|id), data = cs_acc_CAIDE_fmp_post, REML=TRUE)
anova(lmm_CAIDE_cs_post_fmp2, type = "III")
summary(lmm_CAIDE_cs_post_fmp2)$coefficients

sjPlot::tab_model(lmm_CAIDE_cs_post_fmp2, dv.labels = "Task, CAIDE and FMP predicting CS", p.style = "scientific_stars", df.method = "satterthwaite")
```

## Sup8 (Table S7): Mediation Analysis — CAIDE
```{r}
meno_cs_CAIDE_acc <- cleaned_data_df %>%
  select(id, group_var, CAIDE_no_age, cs_rate_easy, cs_rate_hard) %>%
  pivot_longer(!c(id, group_var, CAIDE_no_age), names_to = "task", values_to = "cs_rate") %>%
  mutate(task = str_replace_all(task, "cs_rate_", ""),
         task = factor(task, levels = c("easy", "hard")),
         CAIDE_no_age = factor(CAIDE_no_age, levels = c("0","1","2","3","4","5","6")))

r_helmert_1 <- matrix(c(
  -1/2, 1/2, 0, 0, 0, 0, 0,
  -1/3, -1/3, 2/3, 0, 0, 0, 0,
  -1/4, -1/4, -1/4, 3/4, 0, 0, 0,
  -1/5, -1/5, -1/5, -1/5, 4/5, 0, 0,
  -1/6, -1/6, -1/6, -1/6, -1/6, 5/6, 0,
  -1/7, -1/7, -1/7, -1/7, -1/7, -1/7, 6/7
), ncol = 6)

contrasts(meno_cs_CAIDE_acc$task) <- contr.sum(2)
contrasts(meno_cs_CAIDE_acc$group_var) <- contr.sum(2)
contrasts(meno_cs_CAIDE_acc$CAIDE_no_age) <- r_helmert_1
```

### CAIDE as factor (Table S7)
```{r}
# Required libraries
library(nnet)
library(lme4)
library(sjPlot)

# Path a: group_var → CAIDE_no_age
path_a <- multinom(CAIDE_no_age ~ group_var, data = meno_cs_CAIDE_acc)
summary(path_a)
a_coefs <- as.numeric(summary(path_a)$coefficients[, "group_var1"])

# Path b + c': CAIDE_no_age + group_var → cs_rate
path_b_c <- lmer(cs_rate ~ group_var + CAIDE_no_age + task + (1 | id), 
                 data = meno_cs_CAIDE_acc, REML = TRUE)
summary(path_b_c)

# Coefficients
fixef_coeffs <- fixef(path_b_c)
b_coef <- as.numeric(fixef_coeffs[grep("^CAIDE_no_age", names(fixef_coeffs))])
c_prime_coef <- as.numeric(fixef_coeffs[grep("^group_var", names(fixef_coeffs))])

# Indirect effect (Path a * b)
indirect_effects <- a_coefs * b_coef
total_indirect_effect <- sum(indirect_effects)

# Table S7 output
cat("=== Table S7 Mediation Summary ===\n")
cat("Indirect effects (a * b):\n"); print(indirect_effects)
cat("Total indirect effect:", total_indirect_effect, "\n")
cat("Direct effect (Path c′):", c_prime_coef, "\n")


sjPlot::tab_model(path_b_c,
  dv.labels = "CS Accuracy ~ CAIDE + Menopause + Task",
  p.style = "scientific_stars",
  df.method = "satterthwaite")
```

### Mediation analysis (manual and accounting for task using mixed model)
```{r}
meno_cs_CAIDE_acc$CAIDE_no_age <- as.numeric(meno_cs_CAIDE_acc$CAIDE_no_age)
# Load necessary libraries
library(nnet)   # For multinomial logistic regression
library(lme4)   # For linear mixed-effects models
library(boot)   # For bootstrapping

# 1. Path a: group_var (IV) → CAIDE (Mediator)
# `task` is excluded here because it is not relevant for predicting CAIDE
path_a <- multinom(CAIDE_no_age ~ group_var, data = meno_cs_CAIDE_acc)
summary(path_a)
a_coefs <- as.numeric(summary(path_a)$coefficients[, "group_var1"])

# 2. Path b and Path c': CAIDE (Mediator) + group_var (IV) → cs_rate (DV)
# Include `task` only when predicting cs_rate
path_b_c <- lmer(cs_rate ~ group_var + CAIDE_no_age + task + (1 | id), 
                 data = subset(meno_cs_CAIDE_acc, task == "easy" | task == "hard"),
                 REML = TRUE)
summary(path_b_c)

# Extract coefficients for Path b (CAIDE → cs_rate) and Path c' (group_var → cs_rate)
fixef_coeffs <- fixef(path_b_c)
b_coef <- as.numeric(fixef_coeffs[grep("^CAIDE_no_age", names(fixef_coeffs))])
c_prime_coef <- as.numeric(fixef_coeffs[grep("^group_var", names(fixef_coeffs))])

# 3. Calculate Indirect Effects
indirect_effects <- a_coefs * b_coef  # Element-wise product of Path a and Path b coefficients
total_indirect_effect <- sum(indirect_effects)  # Sum of all indirect effects

# 4. Bootstrapping to Calculate Confidence Intervals
boot_mediation <- function(data, indices) {
  d <- data[indices, ]
  
  # Path a
  path_a <- multinom(CAIDE_no_age ~ group_var, data = d)
  a_coefs <- as.numeric(summary(path_a)$coefficients[, "group_var1"])
  
  # Path b (task included only for cs_rate)
  path_b_c <- lmer(cs_rate ~ group_var + CAIDE_no_age + task + (1 | id), 
                   data = subset(d, task == "easy" | task == "hard"),
                   REML = TRUE)
  fixef_coeffs <- fixef(path_b_c)
  b_coef <- as.numeric(fixef_coeffs[grep("^CAIDE_no_age", names(fixef_coeffs))])
  
  # Indirect Effects
  sum(a_coefs * b_coef)
}

# set.seed(123)
# boot_results <- boot(meno_cs_CAIDE_acc, boot_mediation, R = 1000)
#boot_ci <- boot.ci(boot_results, type = "perc")

# Outputs
print(indirect_effects)          # Indirect effects for each level of CAIDE
print(total_indirect_effect)     # Total indirect effect
print(c_prime_coef)              # Direct effect of group_var on cs_rate
#print(boot_ci)                   # Bootstrapped confidence intervals



confint(path_a) # Confidence intervals for multinomial logistic regression
confint(path_b_c) # Confidence intervals for mixed-effects model



table(meno_cs_CAIDE_acc$group_var, meno_cs_CAIDE_acc$CAIDE_no_age)
```

### Assumption checking
```{r}
# Load necessary libraries
library(car)       # For VIF (multicollinearity check)
library(nnet)      # For multinomial logistic regression
library(lme4)      # For linear mixed-effects models
library(lmtest)    # For heteroscedasticity test
library(ggplot2)   # For visualization
library(boot)      # For bootstrapping
library(dplyr)     # For data manipulation

# Ensure group_var and CAIDE_no_age are factors (categorical variables)
meno_cs_CAIDE_acc <- meno_cs_CAIDE_acc %>%
  mutate(
    group_var = as.factor(group_var),  # Menopause status (Pre vs. Post)
    CAIDE_no_age = as.factor(CAIDE_no_age)  # Categorical CAIDE risk
  )

# === 1. CHECK MULTICOLLINEARITY ===
# Use a Chi-Square test instead of correlation/VIF since both variables are categorical
chi_sq_test <- chisq.test(table(meno_cs_CAIDE_acc$group_var, meno_cs_CAIDE_acc$CAIDE_no_age))

# Print Chi-Square test results
cat("\n=== MULTICOLLINEARITY CHECK (Chi-Square Test) ===\n")
print(chi_sq_test)
table(meno_cs_CAIDE_acc$group_var, meno_cs_CAIDE_acc$CAIDE_no_age)
# === 2. CHECK LINEARITY ===
# Since CAIDE_no_age is categorical, use boxplots instead of scatterplots
ggplot(meno_cs_CAIDE_acc, aes(x = CAIDE_no_age, y = cs_rate)) + 
  geom_boxplot() + ggtitle("Boxplot: CAIDE Category vs. CS Performance")

ggplot(meno_cs_CAIDE_acc, aes(x = group_var, y = cs_rate)) + 
  geom_boxplot() + ggtitle("Boxplot: Menopause Stage vs. CS Performance")

# Residual plot for categorical predictors
fit <- lmer(cs_rate ~ group_var + CAIDE_no_age + task + (1 | id), data = meno_cs_CAIDE_acc, REML = TRUE)
residuals_df <- data.frame(Fitted = fitted(fit), Residuals = resid(fit))

ggplot(residuals_df, aes(x = Fitted, y = Residuals)) + 
  geom_point() + geom_smooth(method = "lm") + ggtitle("Residual Plot")

# === 3. TEST FOR INTERACTION EFFECT (Menopause Stage * CAIDE) ===
interaction_model <- lmer(cs_rate ~ group_var * CAIDE_no_age + task + (1 | id), 
                          data = meno_cs_CAIDE_acc, REML = TRUE)
interaction_summary <- summary(interaction_model)

cat("\n=== TEST FOR INTERACTION ===\n")
print(interaction_summary)

# === 4. CHECK FOR HOMOSCEDASTICITY (Equal Variance of Residuals) ===

library(DHARMa)

# Simulate residuals
sim_res <- simulateResiduals(fittedModel = fit, n = 1000)

# Plot diagnostic residuals
plot(sim_res)

```



## Sup9 (Table S8): Mixed-Effects Models — CAIDE Continuous Predicting CS

### Pre-Menopausal Group
```{r}
cs_acc_CAIDE_pre <- cleaned_data_df %>%
  select(id, group_var, CAIDE, cs_rate_easy, cs_rate_hard) %>%
  filter(group_var == "pre") %>%
  mutate(CAIDE = as.numeric(CAIDE)) %>%
  pivot_longer(!c(id, group_var, CAIDE), names_to = "task", values_to = "cs_rate") %>%
  mutate(task = str_replace_all(task, "cs_rate_", ""),
         task = factor(task, levels = c("easy", "hard")))

contrasts(cs_acc_CAIDE_pre$task) <- contr.sum(2)

lmm_CAIDE_cs_pre <- lmer(cs_rate ~ task + CAIDE + (1 | id), data = cs_acc_CAIDE_pre, REML = TRUE)

anova(lmm_CAIDE_cs_pre, type = "III")
summary(lmm_CAIDE_cs_pre)$coefficients

sjPlot::tab_model(lmm_CAIDE_cs_pre,
  dv.labels = "Pre-Meno:\nTask and CAIDE Risk score predicting CS", p.style = "scientific_stars", df.method = "satterthwaite")

check_collinearity(lmm_CAIDE_cs_pre)
check_outliers(lmm_CAIDE_cs_pre)
```

#### Plot — Pre-Meno
```{r}
effect_plot(lmm_CAIDE_cs_pre, pred = task, plot.points = TRUE, jitter = 0.02,
            line.thickness = 0.7, x.label = "Task Difficulty", y.label = "Correct Source Accuracy") +
  scale_x_discrete(labels = c("Easy", "Hard")) + theme_pubclean()

effect_plot(lmm_CAIDE_cs_pre, pred = CAIDE, plot.points = TRUE, jitter = 0.02,
            line.thickness = 0.7, x.label = "CAIDE", y.label = "Correct Source Accuracy") +
  theme_pubclean()
```

### Post-Menopausal Group
```{r}
cs_acc_CAIDE_post <- cleaned_data_df %>%
    select(id, group_var, CAIDE, cs_rate_easy, cs_rate_hard) %>%
    mutate_at("CAIDE", ~as.numeric(.)) %>%
    filter(group_var=="post" & CAIDE <11) %>%
      pivot_longer(!c(id, group_var, CAIDE), names_to = "task", values_to = "cs_rate") %>%
    mutate(task = str_replace_all(task, "cs_rate_", ""),
         task = factor(task, levels = c("easy", "hard")))


contrasts(cs_acc_CAIDE_post$task) <- contr.sum(2)

lmm_CAIDE_cs_post <- lmer(cs_rate ~ task + CAIDE + (1 | id), data = cs_acc_CAIDE_post, REML = TRUE)

anova(lmm_CAIDE_cs_post, type = "III")
summary(lmm_CAIDE_cs_post)$coefficients

sjPlot::tab_model(lmm_CAIDE_cs_post,
  dv.labels = "Post-Meno:\nTask and CAIDE Risk score predicting CS", p.style = "scientific_stars", df.method = "satterthwaite")

check_collinearity(lmm_CAIDE_cs_post)
check_outliers(lmm_CAIDE_cs_post)
```

#### Plot — Post-Meno
```{r}
effect_plot(lmm_CAIDE_cs_post, pred = task, plot.points = TRUE, jitter = 0.02,
            line.thickness = 0.7, x.label = "Task Difficulty", y.label = "Correct Source Accuracy") +
  scale_x_discrete(labels = c("Easy", "Hard")) + theme_pubclean()

effect_plot(lmm_CAIDE_cs_post, pred = CAIDE, plot.points = TRUE, jitter = 0.02,
            line.thickness = 0.7, x.label = "CAIDE", y.label = "Correct Source Accuracy") +
  theme_pubclean()
```







